---
title: "MA678 homework 05"
subtitle: "Multinomial Regression"
author: "Sky Liu"
date: "Oct. 27, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.width="0.9\\linewidth",dev="pdf",fig.align  = 'center')
pacman::p_load(
  "ggplot2",
  "knitr",
  "arm",
  "data.table",
  "foreign",
  "car",
  "faraway",
  "nnet",
  "reshape2",
  "VGAM",
  "dplyr",
  "nnet")
```

## Multinomial logit: 
Using the individual-level survey data from the 2000 National Election Study (data in folder nes), predict party identification (which is on a 7-point scale) using ideology and demographics with an ordered multinomial logit model.
```{r, echo=FALSE}
nes5200<-read.dta("http://www.stat.columbia.edu/~gelman/arm/examples/nes/nes5200_processed_voters_realideo.dta")
#saveRDS(nes5200,"nes5200.rds")
#nes5200<-readRDS("nes5200.rds")
# logistic regression of vote preference on income
nes5200_dt <- data.table(nes5200)
yr <- 2000
nes5200_dt_s<-nes5200_dt[ year==yr,]
nes5200_dt_s$income <- droplevels(nes5200_dt_s$income)
nes5200_dt_s$partyid7 <- droplevels(nes5200_dt_s$partyid7)

nes5200_dt_s$gender <- factor(nes5200_dt_s$gender, labels=c("male", "female"))
nes5200_dt_s$race <- factor(nes5200_dt_s$race, labels=c("white", "black", "asian", 
                                    "native american", "hispanic"))
nes5200_dt_s$south <- factor(nes5200_dt_s$south)
nes5200_dt_s$ideo <- factor(nes5200_dt_s$ideo, labels=c("liberal", "moderate", "conservative"))

nes_data_comp<-nes5200_dt_s[complete.cases(nes5200_dt_s[,list(partyid7,income,ideo,female,white)])]
nes_data_comp$ideology <- scale(nes_data_comp$ideo_feel,center=TRUE)
```
1. Summarize the parameter estimates numerically and also graphically. 

```{r}
lm1<- polr(partyid7~factor(ideo)+factor(gender),data=nes_data_comp)
summary(lm1)
predx<- expand.grid(ideo = c("moderate","conservative"),gender=c("female","male"))
predy<-predict(lm1,newdata=predx,type = "p")
ggplot(melt(cbind(predx,predy),id.vars = c("gender","ideo")))+
  geom_bar(stat="identity")+aes(x=ideo,y=value, fill=variable)+
  facet_grid(~gender)
```

2. Explain the results from the fitted model.

From the coefficient we can see that gender has no statistical significant impact on party id. 

log odds of not strong democrat = $log(\frac{\pi2+\pi3+\pi4+\pi5+\pi6+\pi7}{\pi1})$

$= \beta_1 moderate+\beta_2 conservative + \beta_3 female-c12$

= 0.5176 + 0.7859 * moderate+ 1.9100 * conservative - 0.3703 * female

log odds of not strong democrat nor weak democrat = $log(\frac{\pi3+\pi4+\pi5+\pi6+\pi7}{\pi1+\pi2})$

$= \beta_1 moderate+\beta_2 conservative + \beta_3 female-c23$

= -0.2747 + 0.7859 * moderate+ 1.9100 * conservative - 0.3703 * female

log odds of not strong democrat, weak democrat nor independent-democrat = $log(\frac{\pi4+\pi5+\pi6+\pi7}{\pi1+\pi2+\pi3})$

$= \beta_1 moderate+\beta_2 conservative + \beta_3 female-c34$

= -1.0234 + 0.7859 * moderate+ 1.9100 * conservative - 0.3703 * female

log odds of not strong democrat, weak democrat, independent-democrat nor independent-independent = $log(\frac{\pi5+\pi6+\pi7}{\pi1+\pi2+\pi3+\pi4})$

$= \beta_1 moderate+\beta_2 conservative + \beta_3 female-c45$

= -1.3914 + 0.7859 * moderate+ 1.9100 * conservative - 0.3703 * female

log odds of weak republican or strong republican = $log(\frac{\pi6+\pi7}{\pi1+\pi2+\pi3+\pi4+\pi5})$

$= \beta_1 moderate+\beta_2 conservative + \beta_3 female-c56$

= -2.1535 + 0.7859 * moderate+ 1.9100 * conservative - 0.3703 * female

log odds of strong republican = $log(\frac{\pi7}{\pi1+\pi2+\pi3+\pi4+\pi5\pi6+})$

$= \beta_1 moderate+\beta_2 conservative + \beta_3 female-c67$

= -3.0906 + 0.7859 * moderate+ 1.9100 * conservative - 0.3703 * female



3. Use a binned residual plot to assess the fit of the model.
```{r}

nesdata <- cbind(partyid=nes_data_comp$partyid7, female=nes_data_comp$female, ideo=nes_data_comp$ideo)
nesdata <- data.frame(na.omit(nesdata))
resid <- model.matrix(~factor(partyid)-1, data=nesdata)-fitted(lm1)
par(mfrow=c(3,2))
for(i in 1:6){
  binnedplot(fitted(lm1)[,i], resid[,i])
}
binnedplot(fitted(lm1)[,7], resid[,7])
```





# High School and Beyond 
The hsb data was collected as a subset of the High School and Beyond study conducted by the National Education Longitudinal Studies program of the National Center for Education Statistics. The variables are gender; race; socioeconomic status; school type; chosen high school program type; scores on reading, writing, math, science, and social studies. We want to determine which factors are related to the choice of the type of program???academic, vocational, or general???that the students pursue in high school. The response is multinomial with three levels.

```{r}
data(hsb)
?hsb
```

1. Fit a trinomial response model with the other relevant variables as predictors (untransformed).
```{r}

lm2 <- multinom(prog~read+write+math+science+race,data = hsb,trace=FALSE,HESS=TRUE) 
summary(lm2)
```

2. For the student with id 99, compute the predicted probabilities of the three possible choices.

```{r}
predict(lm2,newdata=hsb[hsb$id==99,],type="p")
```


# Happiness
Data were collected from 39 students in a University of Chicago MBA class and may be found in the dataset `happy`.
```{r}
library(faraway)
data(happy)
?happy
```

1. Build a model for the level of happiness as a function of the other variables.
```{r}
lm3 <-polr(factor(happy)~money+factor(sex)+factor(love)+factor(work),data=happy) 
summary(lm3)
```

2. Interpret the parameters of your chosen model.
$log\frac{\pi3+...+\pi10}{\pi1+\pi2} = 0.84 + 0.0178 ??? money ??? 1.025 ??? sex_1 + 3.46???love_2 +7.85???love_3 ???1.19???work_2 +0.02???work_3 +1.85???work_4 +0.65???work_5$
Log odds of lonely people who are unsatisfactory with sex, with 0 family income, 
 that has happy index from 3 to 10 over the ones with happy index = 2, is 0.84

3. Predict the happiness distribution for subject whose parents earn $30,000 a year,
who is lonely, not sexually active and has no job.
```{r}
kable(
predict(lm3,newdata=data.frame(love=1,sex=0,work=1,money=30),type="probs"))
```

# newspaper survey on Vietnam War
A student newspaper conducted a survey of student opinions about the Vietnam War in May 1967. Responses were classified by sex, year in the program and one of four opinions. The survey was voluntary. The data may be found in the dataset `uncviet`.  Treat the opinion as the response and the sex and year as predictors. Build a proportional odds model, giving an interpretation to the estimates.

```{r}
data(uncviet)
?uncviet
lm4 <- polr(policy~sex+year,data=uncviet,weights = y,Hess = TRUE) 
summary(lm4)
```

The probability of policy is not A is $exp(1.11???0.65???sexmale+1.18???yearGrad+0.40???yearJunior+0.54???yearSenior+ 0.13 ??? yearSoph)$

A male has opinions B,C or D is 48% (1- exp(-0.65)) lower than a female, holding other variable constant.

A grad student has opinions B,C or D is 224% (exp(1.177)-1) higher than a freshman, holding other variable constant.

A junior student has opinions B,C or D is 49% (exp(0.396)-1) higher than a freshman, holding other variable constant.

A senior student has opinions B,C or D is 72% (exp(0.5444)-1) higher than a freshman, holding other variable constant.

A sophemore student has opinions B,C or D is 14% (exp(1.1315)-1) higher than a freshman, holding other variable constant.


# pneumonoconiosis of coal miners
The pneumo data gives the number of coal miners classified by radiological examination into one of three categories of pneumonoconiosis and by the number of years spent working at the coal face divided into eight categories.

```{r}
library(faraway)
data(pneumo,package="faraway")
head(pneumo)
?pneumo
```

1. Treating the pneumonoconiosis status as response variable as nominal, build a model for predicting the frequency of the three outcomes in terms of length of service and use it to predict the outcome for a miner with 25 years of service.

```{r}
lm6_nominal<-multinom(status~year,weights=Freq,data=pneumo)
summary(lm6_nominal)
pred1<-predict(lm6_nominal,newdata=data.frame(year=25),type = "probs")
```

2. Repeat the analysis with the pneumonoconiosis status being treated as ordinal. 

```{r}
lm6_ornidal<-polr(status~year,weights=Freq,data=pneumo) 
summary(lm6_ornidal)
pred2<-predict(lm6_ornidal,newdata=data.frame(year=25),type = "probs")
```

3.Now treat the response variable as hierarchical with top level indicating whether
the miner has the disease and the second level indicating, given they have the
disease, whether they have a moderate or severe case. 

```{r}
n_pneumo <- pneumo
n_pneumo$status <- as.character(n_pneumo$status)
n_pneumo$status[9:24] <- "abnormal"
n_pneumo$status <- as.factor(n_pneumo$status)
abn_pneumo <- pneumo[-1:-8, ]
lm6_normal <- glm( status ~ year, data = n_pneumo, family = binomial(link = "logit"), weights = Freq)
lm6_abnormal <- glm( status ~ year, data = abn_pneumo, family = binomial(link = "logit"), weights = Freq)

normal<-predict (lm6_normal, newdata=data.frame(year=25), type = "response")
severe<-predict (lm6_abnormal, newdata=data.frame(year=25), type = "response") *(1-predict (lm6_normal, newdata=data.frame(year=25), type = "response"))
mild <- (1-predict (lm6_abnormal, newdata=data.frame(year=25), type = "response")) *(1-predict (lm6_normal, newdata=data.frame(year=25), type = "response"))
pred3<-c(mild, normal, severe)
```

4.  Compare the three analyses.

```{r}
kable(rbind(pred1,pred2,pred3))


new_pneumo<-dcast(pneumo, year ~ status, value.var = "Freq") 
new_pneumo<-new_pneumo %>%mutate(total=apply(new_pneumo[,2:4],1,sum))
new_pneumo[,2:4]<-round(new_pneumo[,2:4]/new_pneumo[,"total"],2) 

pred11<-predict(lm6_nominal,newdata=new_pneumo,type="p")
resid1<-new_pneumo[,2:4]-pred11
par(mfrow=c(1,3))
for(i in 1:3){
  binnedplot(pred11[,i],resid1[,i])
  }

pred22<-predict(lm6_ornidal,newdata=new_pneumo,type="p")
resid2<-new_pneumo[,2:4]-pred22
par(mfrow=c(1,3));for(i in 1:3){binnedplot(pred22[,i],resid2[,i])}




```

From the binnedplot we can see that the second model does not fit well. 

From the probability prediction table we can see that the first and the second model are similar.



# (optional) Multinomial choice models: 

Pardoe and Simonton (2006) fit a discrete choice model to predict winners of the Academy Awards. Their data are in the folder academy.awards.

name  | description
------|----------------------------------------
No    | unique nominee identifier
Year  | movie release year (not ceremony year)
Comp  | identifier for year/category
Name  | short nominee name
PP    | best picture indicator
DD    | best director indicator
MM    | lead actor indicator
FF    | lead actress indicator
Ch    | 1 if win, 2 if lose
Movie | short movie name
Nom   | total oscar nominations
Pic   | picture nom
Dir   | director nom
Aml   | actor male lead nom
Afl   | actor female lead nom
Ams   | actor male supporting nom
Afs   | actor female supporting nom
Scr   | screenplay nom
Cin   | cinematography nom
Art   | art direction nom
Cos   | costume nom
Sco   | score nom
Son   | song nom
Edi   | editing nom
Sou   | sound mixing nom
For   | foreign nom
Anf   | animated feature nom
Eff   | sound editing/visual effects nom
Mak   | makeup nom
Dan   | dance nom
AD    | assistant director nom
PrNl  | previous lead actor nominations
PrWl  | previous lead actor wins
PrNs  | previous supporting actor nominations
PrWs  | previous supporting actor wins
PrN   | total previous actor/director nominations
PrW   | total previous actor/director wins
Gdr   | golden globe drama win
Gmc   | golden globe musical/comedy win
Gd    | golden globe director win
Gm1   | golden globe male lead actor drama win
Gm2   | golden globe male lead actor musical/comedy win
Gf1   | golden globe female lead actor drama win
Gf2   | golden globe female lead actor musical/comedy win
PGA   | producer's guild of america win
DGA   | director's guild of america win
SAM   | screen actor's guild male win
SAF   | screen actor's guild female win
PN    | PP*Nom
PD    | PP*Dir
DN    | DD*Nom
DP    | DD*Pic
DPrN  | DD*PrN
DPrW  | DD*PrW
MN    | MM*Nom
MP    | MM*Pic
MPrN  | MM*PrNl
MPrW  | MM*PrWl
FN    | FF*Nom
FP    | FF*Pic
FPrN  | FF*PrNl
FPrW  | FF*PrWl

```{r, echo=FALSE}
# coefficient for black in 1964 to illustrate nonidentifiability of logistic regression for chap 5 hwk
oscar<-read.csv("http://www.stat.columbia.edu/~gelman/arm/examples/academy.awards/oscars.csv",header=TRUE)
#saveRDS(oscar,"oscar.rds")
#oscar<-readRDS("oscar.rds")
```

1. Fit your own model to these data.

```{r}

```

2. Display the fitted model on a plot that also shows the data.

```{r}

```

3. Make a plot displaying the uncertainty in inferences from the fitted model.

```{r}

```
